The Controlled Choice Design and Paternalism in Pawnshop Borrowing

### 2023-12-17 donde.tex
edit for submission

\vspace{.2in}
\noindent \textbf{Appraising and lending.} Lender P takes gold jewelry as collateral in exchange for a fraction of the value of the piece, in cash. No other collateral and no credit history checks are needed. The transaction takes less than 10 minutes and is conducted at the branch in person between the client and the appraiser (i.e. a teller, see Figure \ref{PawnshopPicture}). The appraiser weighs the gold piece and runs tests on its purity. Based on these she assigns a gold value to the piece, stores it as collateral, and gives 70\% of the gold value of the piece in cash to the client. The borrower signs a 2-page contract with the conditions of the loan and leaves with the cash.

@@

\vspace{.2in}
\noindent \textbf{Many borrowers lose their pawn.} Our context is also one with high borrower default: 43\% of clients lose their pawn in a time span of 230 days from the date of pawning. One potential explanation for high default is that clients are really just knowingly selling their gold piece through a pawn contract on which they intend to default. This appears unlikely for several reasons: (a) clients can easily sell the gold and obtain a higher amount of instant cash at gold-buying stores located close to almost all our pawnshop branches (see Figure \ref{GoldBuyers}), (b) the reported subjective value of the pawn is larger than the appraised value for 83\% of clients, (c) among those that lose their pawn, 29\% paid a positive amount towards its recovery and on average paid 42\% of the value of their loan (see Figure \ref{proxy_naive} in Appendix) --- this can only be rationalized if they expected to recover their pawn, and (d) 72\% of borrowers report a 100\% probability of repaying their loan (and 98\% at least a 50\% chance of repaying) in our baseline survey at the time they take the loan.  %Note that high default could be detrimental from the lender's point of view, since it may reduce the likelihood that the client becomes a return customer. 
%Lender P was in fact explicitly interested in partnering with us to investigate how to reduce loan default, customer satisfaction, and repeat borrowing.

@@

\begin{table}
\caption{Effects on Financial Cost}
\label{main_impact_table}
\begin{center}
\resizebox{1.0\textwidth}{!}{
\scriptsize{\input{./Tables/decomposition_main_te.tex}}
}
\end{center}
\scriptsize{This table shows the treatment effects for our core pecuniary outcomes. Each column is a different regression for different outcomes on an indicator for the forced and choice arms, following specification in equation \ref{basic_reg}. Columns (1) \& (7) analyze our core financial cost measures, while the rest of the columns decompose these into finer components.
A few borrowers take more than one loan on the first day they appear in an experimental branch.  These are treated as different observations. Table \ref{multiple_loans} shows our results are similar when dealing with the multiplicity of loans per borrower. Each regression includes branch and day-of-week FE. Standard errors are clustered at the branch-day level. }

@@

The results from Section \ref{Experiment} show that commitment \emph{works}: clients who were assigned to the forced commitment arm experienced substantially lower financial costs on average.
In spite of this, given the opportunity, only 11\% of borrowers chose commitment. 
If the effect of commitment were homogeneous, this would be enough to conclude that the 89\% who did not choose it would have been financially better-off if they had.
In a world of heterogeneous treatment effects, however, low demand for commitment could still be consistent with borrowers adhering to a standard model of rational choice. 
The borrowers who did not choose commitment could simply be those who don't need it. 
Indeed, we find strong evidence that the effect of forced commitment varies substantially across individuals in our experiment: Appendix \ref{append:chernozhukov} tests and rejects the null hypothesis of homogeneous treatment effects, following the methodology of \cite{chernozhukov2018generic}.
So the question remains: do the 89\% who do not choose commitment know something about their personal situations that we as researchers do not, or are most people in the choice arm making a costly financial mistake? 
In this section we present a series of econometric exercises that sheds light on this question, leveraging unique features of our experimental design. 
Among other results, we show that commitment would lower average financial costs even for the subset of borrowers who choose \emph{not} to commit voluntarily.
To simplify the exposition in this and all sections that follow, we re-define all outcome variables so that \emph{beneficial} treatment effects are \emph{positive}. Using this convention, a positive treatment effect of commitment on financial cost, for example, reflects the average cost \emph{savings} caused by commitment.

@@

\vspace{.2in}
\noindent \textbf{Narrow inputs targeting.} To consider targeting effectiveness we must begin from an individual-specific ground truth, which we take from our estimated conditional average treatment effects (CATEs) $\widehat{\text{ATE}}(X_i)$ from the causal forest described in Section \ref{sec:RF},  which used the ``wide'' set of covariates (wide RF).% and data for all borrowers who replied to at least part of the intake survey. 
We then analyze how well we can predict the losers from paternalism using the ``narrow'' set of non-manipulable covariates listed above (narrow RF) for the same subset of borrowers.
Figure \ref{wide_narrow_forests} shows the relationship between the CATEs estimated from the RF using the ``wide'' and ``narrow'' covariate sets.  We can generate substantially less heterogeneity when using the narrow covariate set, although we still reject the null hypothesis of no treatment effect heterogeneity, following the approach of \cite{chernozhukov2018generic}.\footnote{See \ref{append:chernozhukov} for details of this method.}

>>>>

\vspace{.2in}
\noindent \textbf{Appraising and lending.} Lender P takes gold jewelry as collateral in exchange for a fraction of the value of the piece, in cash. No other collateral and no credit history checks are needed. The transaction takes less than 10 minutes and is conducted at the branch in person between the client and the appraiser (i.e. a teller).
%, see Figure \ref{PawnshopPicture}). 
The appraiser weighs the gold piece and runs tests on its purity. Based on these she assigns a gold value to the piece, stores it as collateral, and gives 70\% of the gold value of the piece in cash to the client. The borrower signs a 2-page contract with the conditions of the loan and leaves with the cash.

@@

\vspace{.2in}
\noindent \textbf{Many borrowers lose their pawn.} Our context is also one with high borrower default: 43\% of clients lose their pawn in a time span of 230 days from the date of pawning. One potential explanation for high default is that clients are really just knowingly selling their gold piece through a pawn contract on which they intend to default. This appears unlikely for several reasons: (a) clients can easily sell the gold and obtain a higher amount of instant cash at gold-buying stores located close to almost all our pawnshop branches,
%(see Figure \ref{GoldBuyers})
(b) the reported subjective value of the pawn is larger than the appraised value for 83\% of clients, (c) among those that lose their pawn, 29\% paid a positive amount towards its recovery and on average paid 42\% of the value of their loan (see Figure \ref{proxy_naive} in Appendix) --- this can only be rationalized if they expected to recover their pawn, and (d) 72\% of borrowers report a 100\% probability of repaying their loan (and 98\% at least a 50\% chance of repaying) in our baseline survey at the time they take the loan.  %Note that high default could be detrimental from the lender's point of view, since it may reduce the likelihood that the client becomes a return customer. 
%Lender P was in fact explicitly interested in partnering with us to investigate how to reduce loan default, customer satisfaction, and repeat borrowing.

@@

\begin{table}
\caption{Effects on Financial Cost}
\label{main_impact_table}
\begin{center}
\resizebox{1.0\textwidth}{!}{
\scriptsize{\input{./Tables/decomposition_main_te.tex}}
}
\end{center}
\scriptsize{This table shows the treatment effects for our core pecuniary outcomes. Each column is a different regression for different outcomes on an indicator for the forced and choice arms, following specification in equation \ref{basic_reg}. Columns (1) \& (7) analyze our core financial cost measures, while the rest of the columns decompose these into finer components.
A few borrowers take more than one loan on the first day they appear in an experimental branch. These are treated as different observations. Additional results, available upon request, show that our results are robust to different ways of handling multiple loans for each borrower. Each regression includes branch and day-of-week FE. Standard errors are clustered at the branch-day level. }
%Table \ref{multiple_loans} shows our results are similar when dealing with the multiplicity of loans per borrower.

@@

The results from Section \ref{Experiment} show that commitment \emph{works}: clients who were assigned to the forced commitment arm experienced substantially lower financial costs on average.
In spite of this, given the opportunity, only 11\% of borrowers chose commitment. 
If the effect of commitment were homogeneous, this would be enough to conclude that the 89\% who did not choose it would have been financially better-off if they had.
In a world of heterogeneous treatment effects, however, low demand for commitment could still be consistent with borrowers adhering to a standard model of rational choice. 
The borrowers who did not choose commitment could simply be those who don't need it. 
Indeed, we find strong evidence that the effect of forced commitment varies substantially across individuals in our experiment: we test and
%Appendix \ref{append:chernozhukov} tests and rejects 
the null hypothesis of homogeneous treatment effects, following the methodology of \cite{chernozhukov2018generic}. (Details available upon request.)
So the question remains: do the 89\% who do not choose commitment know something about their personal situations that we as researchers do not, or are most people in the choice arm making a costly financial mistake? 
In this section we present a series of econometric exercises that sheds light on this question, leveraging unique features of our experimental design. 
Among other results, we show that commitment would lower average financial costs even for the subset of borrowers who choose \emph{not} to commit voluntarily.
To simplify the exposition in this and all sections that follow, we re-define all outcome variables so that \emph{beneficial} treatment effects are \emph{positive}. Using this convention, a positive treatment effect of commitment on financial cost, for example, reflects the average cost \emph{savings} caused by commitment.

@@

\vspace{.2in}
\noindent \textbf{Narrow inputs targeting.} To consider targeting effectiveness we must begin from an individual-specific ground truth, which we take from our estimated conditional average treatment effects (CATEs) $\widehat{\text{ATE}}(X_i)$ from the causal forest described in Section \ref{sec:RF},  which used the ``wide'' set of covariates (wide RF).% and data for all borrowers who replied to at least part of the intake survey. 
We then analyze how well we can predict the losers from paternalism using the ``narrow'' set of non-manipulable covariates listed above (narrow RF) for the same subset of borrowers.
Figure \ref{wide_narrow_forests} shows the relationship between the CATEs estimated from the RF using the ``wide'' and ``narrow'' covariate sets.  We can generate substantially less heterogeneity when using the narrow covariate set, although we still reject the null hypothesis of no treatment effect heterogeneity, following the approach of \cite{chernozhukov2018generic}. \footnote{Details available upon request.}%\footnote{See \ref{append:chernozhukov} for details of this method.}
